Forecasting the 2024 U.S. Presidential Election

### 2024-10-18 paper/paper.qmd
An attempt on building a model

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Tim Chen
  - Steven Li
  - Tommy Fu
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

@@

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
# Set a CRAN mirror
# Set a specific CRAN mirror (this example uses the US CRAN mirror)
install.packages("here")

options(repos = c(CRAN = "https://cran.rstudio.com/"))

# Example of installing a package after setting the mirror
install.packages("tidyverse")

library(tidyverse)
# Install janitor package if not installed
install.packages("janitor")

# Load the janitor package
library(janitor)

install.packages("caret")
install.packages("ROCR")
install.packages("rstanarm")
library(rstanarm)
install.packages("modelsummary")
library(modelsummary)

library(here)

```

@@

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)

# Load the dataset
data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))

```

```{r}
# Build the GLM model
glm_model <- glm(candidate_binary ~ pct + sample_size + methodology + pollscore, 
                 data = data, 
                 family = binomial)

# Summary of the model
summary(glm_model)

```

```{r}
# Make predictions
predictions <- predict(glm_model, newdata = data, type = "response")

# Convert probabilities into binary outcomes (1 for Harris win, 0 for Trump win)
data$predicted_outcome <- ifelse(predictions > 0.5, 1, 0)

# View the prediction results
head(data)

```

```{r}
# Confusion matrix
table(data$candidate_binary, data$predicted_outcome)

# Calculate accuracy
accuracy <- mean(data$candidate_binary == data$predicted_outcome)
accuracy

```

```{r}
# Plot predicted probabilities
ggplot(data, aes(x = pct, y = predictions, color = factor(candidate_binary))) +
  geom_point() +
  labs(title = "Predicted Probability of Harris Winning vs Poll Percentage",
       x = "Polling Percentage", y = "Predicted Probability")

```

@@

```

@@

```
